Non-Vertical Cultural Transmission and the\\ Evolution of Cooperation

### 2020-08-29 ms/main.tex
revised figure captions

\citet{Eshel1982} showed that with assortative meeting, namely, a probability $m$ that individuals interact  within their phenotypic group, cooperation can evolve if $c < b \cdot m$.
Our results highlight another possibility for assortatment, namely, individuals interacting at rate $\alpha$ with their cultural partners, resulting in horizontal transmission.
We show that high levels of assortment  significantly increase the potential for evolution of cooperation. With a high enough $\alpha$, cooperation can increase when initially rare (although it will not fix) even when there is horizontal bias against cooperation ($\alpha > \big(c(1-T_B) + (T_B-T_A)\big) / b T_A$, see Result~\autoref{result:vert_hori})

@@

\citet{woodcock2006significance} showed the importance of non-vertical transmission for the evolution of cooperation. In their model the individuals interact according to standard prisoner dilemma game.
However, they do not link between interaction and transmission. In terms of our model they have choosen the assortative meeting coefficient $\alpha$ to be zero. Our results are more general since it shows the importance of assortative meeting for the evolution of cooperation.

@@

\section*{Figures}
%%% Figure: model illustration
\begin{figure}[thb]
  \centering
  \includegraphics[scale=1]{figure1.pdf}
  \caption{\textbf{Cultural horizontal transmission.} Transmission occurs between interacting partners with probability $\alpha$ (left) or between two random peers with probability $1-\alpha$.}
  \label{fig:horizontal}
\end{figure}

%%% Figure: coexistence map - without oblique transmission
\begin{figure}[htb]
\centering
\includegraphics{coexistence_without_oblique.pdf}
\caption{\textbf{Stable and unstable coexistence between cooperation and defection - without oblique transmission.} The frequency of the cooperative phenotype $A$ among parents in the next generation $\tilde{p}'$ is plotted as a function of the frequency in the current generation $\tilde{p}$.
The line $\tilde{p}'=\tilde{p}$ is shown as a dashed black line. The curves and dashed line intersect at the equilibrium $\tilde{p}^*$ (black circle).
When the curves are above the dashed line ($\tilde{p}<\tilde{p}^*$) the frequency $\tilde{p}$increases towards $\tilde{p}^*$.
When the curves are below the dashed line ($\tilde{p}>\tilde{p}^*$) the frequency $\tilde{p}$ decreases towards $\tilde{p}^*$.
The yellow curve is \textbf{stable coexistence} and has the following parameters values:  $T_A = 0.4$, $T_B = 0.9$, $b = 12$, $c=0.35$, $\alpha = 0.45$. For those values we get that $\gamma_2<c<\gamma_1$ (\autoref{eq:cost_boundaries})
The blue curve is \textbf{unstable coexistence} and has the following parameters values: $T_A = 0.5$, $T_B = 0.1$, $b = 1.3$, $c=0.904$, $\alpha = 0.4$. For those values we can that $\gamma_1<c<\gamma_2$}
\label{fig:coexistence_without_oblique}
\end{figure}

@@

    \includegraphics[scale=0.75]{boundiries_withAlpha_y_axis.pdf}
    \caption{\textbf{Evolution of cooperation under vertical and horizontal cultural transmission.} 
The figure shows the global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or defection depending on the initial conditions - unstable coexistence (blue), and stable coexistence of cooperation and defection (yellow).
In all cases the horizontal bias ($T_A-T_B$) is on the x-axis.
(\textbf{a-b})~the cost of cooperation $c$ is on the y-axis; the cost boundaries $\gamma_1$ and $\gamma_2$ (\autoref{eq:cost_boundaries}) are the solid and dashed lines. 
(\textbf{c})~the assortment $\alpha$ is on the y-axis; the assortment boundaries $a_1$ and $a_2$ (\autoref{eq:boundries_assortative_meeting}) are the solid and dashed lines.
Here, $b=1.3$, $T_A=0.4$. In figure (\textbf{a})~$\alpha = 0.7$ and in figure (\textbf{b})~$\alpha = 0$. In figure (\textbf{c})~$c = 0.35$.
  % TODO missing text descriptions of conditions on (3)
  	}
    \label{fig:results2}
\end{figure}

%%% Figure: coexistence map - without oblique transmission
\begin{figure}[htb]
  \centering
  \includegraphics{coexistence_with_oblique.pdf}
  \caption{\textbf{Stable coexistence between cooperation and defection - with oblique transmission.} The frequency of the cooperative phenotype $A$ among juveniles in the next generation $\hat{p}'$ is plotted as a function of the frequency in the current generation $\hat{p}$.
  The line $\hat{p}'=\hat{p}$ is shown as a dashed black line. The curve and dashed line intersect at the equilibrium $\hat{p}^*$ (black circle).
  When the curves are above the dashed line ($\hat{p}<\hat{p}^*$) the frequency $\hat{p}$increases towards $\hat{p}^*$.
  When the curves are below the dashed line ($\hat{p}>\hat{p}^*$) the frequency $\hat{p}$ decreases towards $\hat{p}^*$.
  The curve is \textbf{stable coexistence} and has the following parameters values:  $T_A = 0.4$, $T_B = 0.9$, $b = 20$, $c=0.1$, $\alpha = 1$, $v=0.4$. For those values we get that $0<\beta_3<\beta_1$ (\autoref{eq:polynomial_coefficients})}
  \label{fig:coexistence_with_oblique}
  \end{figure}

>>>>

\citet{Eshel1982} showed that with assortative meeting, namely, a probability $m$ that individuals interact  within their phenotypic group, cooperation can evolve if $c < b \cdot m$.
Our results highlight another possibility for assortment, namely, individuals interacting at rate $\alpha$ with their cultural partners, resulting in horizontal transmission.
We show that high levels of assortment  significantly increase the potential for evolution of cooperation. With a high enough $\alpha$, cooperation can increase when initially rare (although it will not fix) even when there is horizontal bias against cooperation ($\alpha > \big(c(1-T_B) + (T_B-T_A)\big) / b T_A$, see Result~\autoref{result:vert_hori})

@@

\citet{woodcock2006significance} showed the importance of non-vertical transmission for the evolution of cooperation. In their model the individuals interact according to standard prisoner dilemma game.
However, they do not link between interaction and transmission. In terms of our model they have chosen the assortative meeting coefficient $\alpha$ to be zero. Our results are more general since it shows the importance of assortative meeting for the evolution of cooperation.

@@

\section*{Figures}
%%% Figure: model illustration
\begin{figure}[thb]
  \centering
  \includegraphics[scale=1]{figure1.pdf}
  \caption{\textbf{Cultural horizontal transmission with assortment.} Transmission occurs between interacting partners with probability $\alpha$ (left) or between two random peers with probability $1-\alpha$.}
  \label{fig:horizontal}
\end{figure}

%%% Figure: coexistence map - without oblique transmission
\begin{figure}[htb]
\centering
\includegraphics{coexistence_without_oblique.pdf}
\caption{\textbf{Stable and unstable coexistence between cooperation and defection without oblique transmission.}
  The curves show the frequency of the cooperative phenotype $A$ among parents in the next generation $\tilde{p}'$ vs. the current generation $\tilde{p}$ (\autoref{eq:nextgen_parents_vertical_only}).
  The dashed black line is $\tilde{p}'=\tilde{p}$.
  The curves and the dashed line intersect at the polymorphic equilibrium $\tilde{p}^*$ (black circle).
  When the curves are above the dashed line, $\tilde'{p}>\tilde{p}$, then $\tilde{p}$ increase.
  When the curves are below the dashed line, $\tilde'{p}<\tilde{p}$, then $\tilde{p}$ decreases.
  The yellow curve, for which the polymorphic equilibrium is stable, is given by $T_A = 0.4$, $T_B = 0.9$, $b = 12$, $c=0.35$, and $\alpha = 0.45$, which give $\gamma_2<c<\gamma_1$ (\autoref{eq:cost_boundaries})
  The blue curve, for which the equilibrium is unstable, is given by $T_A = 0.5$, $T_B = 0.1$, $b = 1.3$, $c=0.904$, and $\alpha = 0.4$, which give $\gamma_1<c<\gamma_2$.
  In both cases there is no oblique transmission, $v=1$; see \autoref{fig:coexistence_with_oblique} for $v<1$.
  }
\label{fig:coexistence_without_oblique}
\end{figure}

@@

    \includegraphics[scale=0.75]{boundiries_withAlpha_y_axis.pdf}
    \caption{\textbf{Evolution of cooperation under vertical and horizontal cultural transmission.} 
The figure shows the global fixation of cooperation (green), global fixation of defection (red), fixation of either cooperation or defection depending on the initial conditions, i.e. unstable coexistence (blue), and stable coexistence of cooperation and defection (yellow).
In all cases the horizontal bias ($T_A-T_B$) is on the x-axis.
(\textbf{a-b})~the cost of cooperation $c$ is on the y-axis; the cost boundaries $\gamma_1$ and $\gamma_2$ (\autoref{eq:cost_boundaries}) are the solid and dashed lines. 
(\textbf{c})~the assortment $\alpha$ is on the y-axis; the assortment boundaries $a_1$ and $a_2$ (\autoref{eq:boundries_assortative_meeting}) are the solid and dashed lines.
Here, $b=1.3$, $T_A=0.4$. (\textbf{a})~$\alpha = 0.7$.(\textbf{b})~$\alpha = 0$. (\textbf{c})~$c = 0.35$.
  % TODO missing text descriptions of conditions on (3)
  	}
    \label{fig:results2}
\end{figure}

%%% Figure: coexistence map - without oblique transmission
\begin{figure}[htb]
  \centering
  \includegraphics{coexistence_with_oblique.pdf}
  \caption{\textbf{Stable coexistence between cooperation and defection with oblique transmission.}
  The curve shows the frequency of the cooperative phenotype $A$ among juveniles in the next generation $\hat{p}'$ vs. the current generation $\hat{p}$ (\autoref{eq:horizontal}).
  The dashed black line is $\hat{p}'=\hat{p}$.
  The curve and the dashed line intersect at the stable equilibrium $\hat{p}^*$ (black circle).
  When $\hat{p} < \hat{p}^*$ then the curve is above the dashed line, $\hat{p}' > \hat{p}$, and $\hat{p}$ increases towards $\hat{p}^*$.
  When $\hat{p} > \hat{p}^*$ then the curve is below the dashed line, $\hat{p}' < \hat{p}$, and $\hat{p}$ decreases towards $\hat{p}^*$.
  Here,  $T_A = 0.4$, $T_B = 0.9$, $b = 20$, $c=0.1$, $\alpha = 1$, and $v=0.4$, which give $0<\beta_3<\beta_1$ (\autoref{eq:polynomial_coefficients}).
  }
  \label{fig:coexistence_with_oblique}
  \end{figure}
